---
title: "CMR Results"
author: "Limoilou"
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: cmr.bib
csl: canadian-journal-of-zoology.csl
fontsize: 12pt
fontfamily: times
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, verbose=FALSE, cache=TRUE)
# knitr::opts_knit$set(root.dir = '~/Users/LimoilouARenaud/Documents/PostDocI/Projects/cmrSeal')
```

```{r include=FALSE}
library(knitr)
library(readxl)
library(magrittr) # problems with tidyverse. Installed this instead
library(dplyr)
library(coda)
library(bookdown)
library(ggplot2)
library(ggthemes)
library(pander)
library(cowplot)
library(lubridate)
library(boot)
library(purrr)
library(kableExtra)
library(here)
```

# model selection at bic - abundance 
```{r}
load(here::here("output.nosync/data/2022-12-07_modelSelection.RData"))

aicT[1,2]<-"$p(.)\\phi(.)b(t)$"# could put math in markdown - tex though
aicT[2,2]<-"$p(t)\\phi(.)b(t)$"
aicT[3,2]<-"$p(t)\\phi(t)b(t)$"
aicT[4,2]<-"$p(.)\\phi(t)b(t)$"
aicT[5,2]<-"$p(1|t)\\phi(1|t)b(t)$"
#aicT[6,2]<-"$p()\\phi()\\gamma(t)$" # will have to ad 6 and 7

aicT=aicT %>% mutate(across(where(is.numeric), round, 2))

long_N_dt_b <- cbind(aicb1[,c(1,3)], aicb2[3], aicb3[3],aicb4[3], aicb5[3])
names(long_N_dt_b)<-c("year", "model1", "model2", "model3", "model4", "model5")
long_N_dt_b <- long_N_dt_b %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(long_N_dt_b,caption = "Table S4 Selected model per year, Bic, Qc, Canada.", digits=0) %>% kable_styling()# %>%save_kable(here("tableS4_modeSelBic.html"))


bestmod <- apply(long_N_dt_b[,-1],1,which.min)
tmp <-cbind(long_N_dt_b$year,colnames(long_N_dt_b)[bestmod+1]) # m3


```


# model selection at metis  - abundance 
```{r import RData from R}
aicTM[1,2]<-"$p(.)\\phi(.)b(t)$"# could put math in markdown - tex though
aicTM[2,2]<-"$p(t)\\phi(.)b(t)$"
aicTM[3,2]<-"$p(t)\\phi(t)b(t)$"
aicTM[4,2]<-"$p(.)\\phi(t)b(t)$"
aicTM[5,2]<-"$p(1|t)\\phi(1|t)b(t)$"
#aicTM[6,2]<-"$p()\\phi()\\gamma(t)$"

aicTM=aicTM %>% mutate(across(where(is.numeric), round, 2))

#knitr::kable(aicTM, caption = "Average WAIC per model over years, Métis, Qc, Canada.", align = "llr")

long_N_dt_m <- cbind(aicm1[,c(1,3)], aicm2[3], aicm3[3],aicm4[3], aicm5[3])
names(long_N_dt_m)<-c("year", "model1", "model2", "model3", "model4", "model5")
long_N_dt_m <- long_N_dt_m %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(long_N_dt_m,caption = "Table S4. Selected model per year, Metis, Qc, Canada.", digits=0) %>% kable_styling()#  %>%save_kable(here("tableS4_modeSelMetis.html"))

bestmod <- apply(long_N_dt_m[,-1],1,which.min)
tmp <-cbind(long_N_dt_m$year,colnames(long_N_dt_m)[bestmod+1]) #m3
```


# supp table 1 with modSel
```{r JS model description, echo=FALSE}
jsmodels <- aicT[1:5, 1:2]
jsmodels$details <-NA

names(jsmodels) <- c('model', 'notation', 'description')
jsmodels[1,3]<-"A constant detection probability p and survival probability φ, across all individuals and occasions."
jsmodels[2,3]<-"A fixed time effect to account for within-season variation in p and a constant φ over the season. As pups grow, they are less easily captured (Dubé et al. 2003)."
jsmodels[3,3]<-"A fixed time effect on detection probability p, and a fixed time effect on survival probability φ."
jsmodels[4,3]<-"A time effect on φ but a constant capture probability, p."
jsmodels[5,3]<-"A random effect of time on both p and φ if individuals do not have the same capture or survival probability across the season because of local environmental conditions affecting swimming behavior, for example."

knitr::kable(jsmodels,caption = "Description of the Jolly-Seber models for the estimation of harbour seal pup abundance in the St. Lawrence Estuary, Quebec, Canada, 1998-2019. p: capture probability, b: recruitment, φ: survival probability, (.): a constant effect over sampling occasions, (t): a fixed time effects over sampling occasions, (1|t) : random time effects (in survival or capture probabilities). All models assumed fixed time effects in entry probabilities, b.", "simple")
```

# supp table 2

```{r JS model selection for both sites, echo=FALSE}
names <- c(
  ' ', "$p(.)\\phi(.)b(t)$","$p(t)\\phi(.)b(t)$", "$p(t)\\phi(t)b(t)$", "$p(.)\\phi(t)b(t)$", "$p(1|t)\\phi(1|t)b(t)$"
)
long_N_dt_b <- rbind(names, long_N_dt_b)
long_N_dt_m <- rbind(names, long_N_dt_m)

names2 <- c('','Bic Island',  rep('', 4), 'Metis', rep('', 4))
modSel_JS <- cbind(long_N_dt_b, long_N_dt_m[, 2:6])
modSel_JS <- rbind(names2, modSel_JS)

knitr::kable(modSel_JS,caption = "Jolly-Seber model selection for the estimation of pup abundance in a harbour seal population studied at Bic Island and Métis, Quebec, Canada, 1998-2019. Each year was modeled independently thus yielding a WAIC per year; the best model was selected based on number of years it yielded the lowest WAIC. p: capture probability, b: recruitment, φ: survival probability, (.): a constant effect over sampling occasions, (t): a fixed time effects over sampling occasions, (1|t) : temporal random effects (in survival and capture probabilities).", "simple")
```



# model selection at bic - survival 


```{r CJS model description, echo=FALSE}
# cjsmodels <- waic.table[1:6, 1:2]
# cjsmodels$details <-NA
# 
# names(cjsmodels) <- c('model', 'notation', 'description')
# 
# 
# cjsmodels[1,3]<-"Constant detection probability p and pre-weaning survival probability φ, across all pups and occasions."
# cjsmodels[2,3]<-"A fixed time effect to account for within-season variation in p, and a constant φ over the season."
# cjsmodels[3,3]<-"A fixed time effect on detection probability p, and a random time effect on survival probability φ."
# cjsmodels[4,3]<-"A constant capture probability p, and a random time effect on survival probability φ if pups do not have the same survival probability across the season."
# cjsmodels[5,3]<-"A random individual effect on p to model capture heterogeneity, and a constant φ across the season."
# cjsmodels[6,3]<-"A random individual effect on p to model capture heterogeneity, and a random time effect on φ if pups do not have the same survival probability across the season."
# 
# knitr::kable(cjsmodels,caption = "Description of the Cormack-Jolly-Seber models for the estimation of harbour seal pre-weaning survival in the St. Lawrence Estuary, Quebec, Canada, 1998-2019. p: capture probability, φ: pre-weaning survival probability at 30 days, (.): a constant effect over sampling occasions, (t): a fixed time effects over sampling occasions, (1|t) : random time effects (in survival or capture probabilities). All models assumed fixed time effects in entry probabilities, b.", "simple")
```



```{r best CJS model}
# recreate waic.table since no mean  anymore 
waic.table<-data.frame(model = c(1:12), description=NA)

# # give proper names to the models
# waic.table[1,2]<-"$p(.)\\phi(.)$"# could put math in markdown - tex though
# waic.table[2,2]<-"$p(t)\\phi(.)$"
# waic.table[3,2]<-"$p(t)\\phi(1|t)$"
# waic.table[4,2]<-"$p(.)\\phi(1|t)$"
# waic.table[5,2]<-"$p(1|i)\\phi(.)$"
# waic.table[6,2]<-"$p(1|i)\\phi(1|t)$" # will have to ad 6 and 7
# waic.table[7,2]<-"$p(.)\\phi(.)$"
# waic.table[8,2]<-"$p(t)\\phi(.)$"
# waic.table[9,2]<-"$p(t)\\phi(1|t)$"
# waic.table[10,2]<-"$p(.)\\phi(1|t)$"
# waic.table[11,2]<-"$p(1|i)\\phi(.)$"
# waic.table[12,2]<-"$p(1|i)\\phi(1|t)$"

waic.table <- waic.table %>% mutate(across(where(is.numeric), round, 2))


# determine which is best consdiereing year 2008 at bic shouldnt be considered 
long_Surv_dt_b <- cbind(WAIC.b.1[,c(2,1)], WAIC.b.2[1], WAIC.b.3[1],WAIC.b.4[1], WAIC.b.5[1], WAIC.b.6[1])
names(long_Surv_dt_b)<-c("year", "model1", "model2", "model3", "model4", "model5", 'model6')
long_Surv_dt_b <- long_Surv_dt_b %>% mutate(across(where(is.numeric), round, 0))
knitr::kable(long_Surv_dt_b,caption = "Table S5. Selected model per year, Bic, Qc, Canada.",  digits=0) %>% kable_styling()#%>%save_kable(here("tableS5_modeSelBic.html"))

# immportant - to do
# long_Surv_dt_b=long_Surv_dt_b %>% filter(year!=2003&year!=2008&year!=2009)
long_Surv_dt_b=long_Surv_dt_b %>% filter(year!=2002&year!=2008) # m3 and m2 equal but m3 has time variation in surv


bestmod <- apply(long_Surv_dt_b[,-1],1,which.min) # model 3 - 6 out of 15 times but need to remove bad years - 2008 here 
tmp <-cbind(long_Surv_dt_b$year,colnames(long_Surv_dt_b)[bestmod+1])
knitr::kable(tmp,caption = "Selected model per year, Bic, Qc, Canada.", "simple")



# model 3 and 2 equally chosen when excluding the 2 years with low convergence according to gelman but model 2 has a constant surv over occasions which is useless 

```


# model selection at metis - survival 


ATTENTION 

make sure that years with low convergence are removed from calculations. 


```{r}
# determine which is best at metis - a few years did not converge including 2008

long_Surv_dt_m <- cbind(WAIC.m.1[,c(2,1)], WAIC.m.2[1], WAIC.m.3[1],WAIC.m.4[1], WAIC.m.5[1], WAIC.m.6[1])
names(long_Surv_dt_m)<-c("year", "model1", "model2", "model3", "model4", "model5", "model6")
long_Surv_dt_m <- long_Surv_dt_m %>% mutate(across(where(is.numeric), round,0))
knitr::kable(long_Surv_dt_m,caption = "Table S5. Selected model per year, Metis, Qc, Canada.", digits=0) %>% kable_styling()#  %>%save_kable(here("tableS5_modeSelMetis.html"))

# with new runs - worst years in terms of gelman
#long_Surv_dt_m=long_Surv_dt_m %>% filter(year!=1999&year!=2002&year!=2009)
long_Surv_dt_m=long_Surv_dt_m %>% filter(year!=2000&&year!=2002&year!=2013)

bestmod <- apply(long_Surv_dt_m[,-1],1,which.min)
tmp <-cbind(long_Surv_dt_m$year,colnames(long_Surv_dt_m)[bestmod+1])
knitr::kable(tmp,caption = "Selected model per year, Métis, Qc, Canada.", "simple")

# no clear best model - model 3 chosen twice, model 2 4 times, and 1 chosen 3 times 
# the year we exclude makes a big differnece

```



```{r ts4, echo=FALSE, verbose=FALSE}
names <- c(' ', "$p(.)\\phi(.)$", "$p(t)\\phi(.)$", "$p(t)\\phi(1|t)$", "$p(.)\\phi(1|t)$", "$p(1|i)\\phi(.)$", "$p(1|i)\\phi(1|t)$" )
long_Surv_dt_b <- rbind(names, long_Surv_dt_b)
knitr::kable(long_Surv_dt_b,caption = "Cormack-Jolly-Seber model selection for the estimation of pre-weaning survival in a harbour seal pups, Bic Island, Quebec, Canada. Each year was modeled independently thus yielding a WAIC per year; the best model was selected based on number of years it yielded the lowest WAIC. p: capture probability, phi : survival probability, (.) : a constant effect over sampling occasions, (t) : a fixed time effects over sampling occasions, (1|t) : temporal random effects (in survival and capture probabilities), (1|i) : individual random effects (capture probabilities).", "simple")
```


```{r ts5, echo=FALSE, verbose=FALSE}
long_Surv_dt_m <- rbind(names, long_Surv_dt_m)
knitr::kable(long_Surv_dt_m,caption = "Cormack-Jolly-Seber model selection for the estimation of pre-weaning survival in a harbour seal pups, Métis, Quebec, Canada. Each year was modeled independently thus yielding a WAIC per year; the best model was selected based on number of years it yielded the lowest WAIC. p: capture probability, phi : survival probability, (.) : a constant effect over sampling occasions, (t) : a fixed time effects over sampling occasions, (1|t) : temporal random effects (in survival and capture probabilities), (1|i) : individual random effects (capture probabilities).", "simple")
```

# sample size 

New data frame - ATtenton one df for JS, one for CJS

```{r sample size}
rm(list =ls())

library(here)


load(here("data/mine/2022-12-05_dataCJS_pup35.RData"))
years


data <- pup.data35
nlevels(droplevels(unique(data$myID)))
length(unique(data$myID)) #1371
length(unique(data$date)) #  308
table(data$sex)
# female   male     NA
#   1120   1252     8 #
t = table(data$myID)
#View(t)
mean(t)
table(t)

#   0   1   2   3   4   5   6   7   8
#  13 756 379 138  63  17  12   5   1 changed 2022 09 -09 
sum(t)
756/1371# 0.5521517
(379 +138 + 63 + 17 +12+   5 +  1 )/1371 #  0.4485777



#   1   2   3   4   5   6   7   8 
# 758 378 138  63  17  12   5   1


sum(t)
758/1373#  0.5520757
(378 +138 + 63 + 17 +12+   5 +  1 )/1372 #  0.4475219



# make table 1
data$myID <- droplevels(data$myID)
t1 <- table(data$myID)
# View(t1)
tmp=data[data$myID %in% names(t1)[t1>1],]
#View(tmp)
nlevels(droplevels(unique(tmp$myID))) # 615 individus 615/1371
t2=table(tmp$myID)
#View(t2)

s.sizes.b = data %>%
  filter(mySite=="bic") %>%
  group_by(year) %>%
  mutate(date =ymd(date)) %>%
           summarise(min = min(date),
                     max= max(date),
                     unique.ID = length(unique(myID)),
                     nb.male=length(na.omit(unique(ifelse(sex=='male',myID,NA)))),
            nb.female=length(na.omit(unique(ifelse(sex=='female',myID,NA))))
)

recap.b=pup.data35 %>%
  filter(mySite=="bic") %>%
  group_by(year) %>%
  droplevels() %>%
  summarise(unique.ID = length(unique(myID)))
#write.csv2(recap.b, "output/data/recapture_b.csv")

# add prop of recapt for table 1
t1 <- table(pup.data35$myID)
tmp=pup.data35[pup.data35$myID %in% names(t1)[t1>1],]

t2=tmp %>%
  filter(mySite=="bic") %>% group_by(year) %>%
  droplevels %>%
  summarise(recapt = length(unique(myID)))
recap.b=cbind(recap.b, t2[,2])
recap.b=recap.b%>% mutate(prop=recapt/unique.ID)

s.sizes.b <- cbind(s.sizes.b, recap.b[,4])


s.sizes.m = data %>%
  filter(mySite=="metis") %>%
  group_by(year) %>%
  mutate(date =ymd(date)) %>%
           summarise(min = min(date),
                     max= max(date),
                     unique.ID = length(unique(myID)),
                     nb.male=length(na.omit(unique(ifelse(sex=='male',myID,NA)))),
            nb.female=length(na.omit(unique(ifelse(sex=='female',myID,NA))))
)

recap.m=pup.data35 %>%
  filter(mySite=="metis") %>%
  group_by(year,.drop = FALSE) %>%
  #droplevels() %>%
  summarise(unique.ID = length(unique(myID)))
#write.csv2(recap.m, "output/data/recapture_m.csv")

t1 <- table(pup.data35$myID)
tmp=pup.data35[pup.data35$myID %in% names(t1)[t1>1],]

t2=tmp %>%
  filter(mySite=="metis") %>% group_by(year,.drop = FALSE) %>%
  #droplevels() %>%
  summarise(recapt = length(unique(myID)))

recap.m=cbind(recap.m, t2[,2])
recap.m=recap.m%>% mutate(prop=recapt/unique.ID)

s.sizes.m <- cbind(s.sizes.m, recap.m[,4])


# quantify effort - number of days with at least one capture 

effortB=bicData35. %>% group_by(year) %>% summarise(effort = length(unique(date)))
effortM=metisData35. %>% group_by(year) %>% summarise(effort = length(unique(date)))

effortB=bicData35. %>% group_by(year) %>% summarise(effort = length(unique(date)))



```

# CMR results 


```{r}
#load(here("output.nosync/data/2022-12-09_resultsCMR.RData"))
load(here("output.nosync/data/2022-12-23_resultsCMR.RData"))

```


# revisions by Mike - annual trends in N or surv

```{r}
#  annual trend in abundance at BIC?
summary(lm(N_bic.3_summa$Nsuper_mean~as.numeric(N_bic.3_summa$yr))) #NS
# Coefficients:
#                                Estimate Std. Error t value Pr(>|t|)    
# (Intercept)                  -13925.355   2599.525  -5.357 0.000101 ***
# as.numeric(N_bic.3_summa$yr)      6.993      1.294   5.402 9.33e-05 ***
plot(N_bic.3_summa$Nsuper_mean~as.numeric(N_bic.3_summa$yr))

confint(lm(N_bic.3_summa$Nsuper_mean~as.numeric(N_bic.3_summa$yr)))

#  2.5 %       97.5 %
# (Intercept)                  -19500.780690 -8349.928436
# as.numeric(N_bic.3_summa$yr)      4.216104     9.768942


# metis?
summary(lm(N_metis.3_summa$Nsuper_mean~as.numeric(N_metis.3_summa$yr))) #NS

# Coefficients:
#                                  Estimate Std. Error t value Pr(>|t|)
# (Intercept)                    -1947.6906  1282.8368  -1.518    0.151
# as.numeric(N_metis.3_summa$yr)     0.9941     0.6388   1.556    0.142
plot(N_metis.3_summa$Nsuper_mean~as.numeric(N_metis.3_summa$yr))

confint(lm(N_metis.3_summa$Nsuper_mean~as.numeric(N_metis.3_summa$yr))) #
# 2.5 %     97.5 %
# (Intercept)                    -4699.1018212 803.720706
# as.numeric(N_metis.3_summa$yr)    -0.3759855   2.364279
```

Temporal trends in survival 

```{r}
# at bic
summary(lm(weanSurvOut_b_summa$weanSurv_med~as.numeric(weanSurvOut_b_summa$yr)))
# as.numeric(weanSurvOut_b_summa$yr) -0.006659   0.010486  -0.635    0.536


# at metis 

summary(lm(weanSurvOut_m_summa$weanSurv_med~as.numeric(weanSurvOut_m_summa$yr)))
# as.numeric(weanSurvOut_m_summa$yr) -0.023309   0.006335  -3.680  0.00248 **


```

# revisions by Mike - effect of capture effort on surv 

```{r}

# added 2022 11 11 by Mike - is interannual diff explained by capture effort?
hist(weanSurvOut_b_summa$weanSurv_mean)
hist(weanSurvOut_m_summa$weanSurv_mean)

weanSurvOut_b_summa$effort <- effortB$effort
weanSurvOut_m_summa$effort <- effortM$effort

summary(lm(weanSurvOut_b_summa$weanSurv_mean~weanSurvOut_b_summa$effort))
# weanSurvOut_b_summa$effort  0.03613    0.01391   2.597   0.0211 *
m1=lm(weanSurvOut_b_summa$weanSurv_mean~weanSurvOut_b_summa$effort)
qqnorm(resid(m1));qqline(resid(m1)) #
summary(glm(weanSurvOut_b_summa$weanSurv_mean~weanSurvOut_b_summa$effort, family = 'binomial'))
# weanSurvOut_b_summa$effort   0.1546     0.1411   1.096    0.273
summary(lm(weanSurvOut_b_summa$weanSurv_med~weanSurvOut_b_summa$effort))
# weanSurvOut_b_summa$effort  0.03734    0.01596   2.340   0.0346 *
summary(glm(weanSurvOut_b_summa$weanSurv_med~weanSurvOut_b_summa$effort, family = 'binomial'))
# weanSurvOut_b_summa$effort   0.1607     0.1421   1.131    0.258

plot(weanSurvOut_b_summa$weanSurv_mean~weanSurvOut_b_summa$effort)
plot(weanSurvOut_b_summa$weanSurv_med~weanSurvOut_b_summa$effort)

summary(lm(weanSurvOut_m_summa$weanSurv_mean~weanSurvOut_m_summa$effort))
# weanSurvOut_m_summa$effort  0.021645   0.008406   2.575    0.022 *
summary(lm(weanSurvOut_m_summa$weanSurv_med~weanSurvOut_m_summa$effort))
# weanSurvOut_m_summa$effort  0.02428    0.01003   2.422   0.0296 *
    
plot(weanSurvOut_m_summa$weanSurv_mean~weanSurvOut_m_summa$effort)
plot(weanSurvOut_m_summa$weanSurv_med~weanSurvOut_m_summa$effort)

m2=glm(weanSurvOut_m_summa$weanSurv_mean~weanSurvOut_m_summa$effort, family = binomial(link = 'logit'))
# weanSurvOut_m_summa$effort   0.1332     0.1315   1.013   0.3111  


# in abundance too?
N_bic.3_summa$effort <- effortB$effort
summary(lm(N_bic.3_summa$Nsuper_mean~N_bic.3_summa$effort)) #NS


N_metis.3_summa$effort <- effortM$effort
summary(lm(N_metis.3_summa$Nsuper_mean~N_metis.3_summa$effort)) #NS


```

## any effect of seals or environmental variables 

Correlations between grey seal abundance and preweaning survival
```{r}
# combine grey seal abundance to df of surv
totalAbundance_gs <- read.csv("~/Documents/PostDocI/Projects/cmrSeal/data/mine/ensembleTotalAbundance.csv")
gulfAbundance_gs <- read.csv("~/Documents/PostDocI/Projects/cmrSeal/data/mine/ensembleGulfAbundance.csv")
climate_PG <- read_excel("~/Documents/PostDocI/Projects/cmrSeal/data/mine/climate_PG.xlsx", sheet = "Feuil1")
colnames(climate_PG) <- c('year', 'surf_anomaly', 'intermediate_anomaly')

# add total abun bic
tmp_bic <- merge(weanSurvOut_b_summa, totalAbundance_gs[, c('year','median')], by.x='yr', by.y='year') %>% rename(med_tot=median)

# add gulf abun bic
tmp2_bic <- merge(tmp_bic, gulfAbundance_gs[, c('year','median')], by.x='yr', by.y='year')%>% rename(med_gulf=median)

tmp2_bic <- merge(tmp2_bic, climate_PG, by.x='yr', by.y='year') %>% filter(yr!=2002&yr!=2008)


summary(lm(tmp2_bic$weanSurv_mean~tmp2_bic$med_tot)) #NS tmp2_bic$med_tot -0.0007804  0.0010866  -0.718   0.4863  
 
summary(lm(tmp2_bic$weanSurv_mean~tmp2_bic$med_gulf)) # NS tmp2_bic$med_gulf -0.01028    0.01076  -0.956   0.3581  

summary(lm(tmp2_bic$weanSurv_mean~tmp2_bic$surf_anomaly)) #NS
summary(lm(tmp2_bic$weanSurv_mean~tmp2_bic$intermediate_anomaly)) #NS tmp2_bic$intermediate_anomaly  0.06746    0.07474   0.903    0.384    

cor.test(tmp2_bic$weanSurv_mean,tmp2_bic$surf_anomaly, method='pearson') # 0.3895855 p-value = 0.1685
cor.test(tmp2_bic$weanSurv_mean,tmp2_bic$intermediate_anomaly, method='pearson') # 0.2521395 p-value = .3845



      
# add total abun metis
tmp_metis <- merge(weanSurvOut_m_summa, totalAbundance_gs[, c('year','median')], by.x='yr', by.y='year') %>% rename(med_tot=median)

# add gulf abun metis
tmp2_metis <- merge(tmp_metis, gulfAbundance_gs[, c('year','median')], by.x='yr', by.y='year')%>% rename(med_gulf=median)


# metis
summary(lm(weanSurvOut_m_summa$weanSurv_mean~weanSurvOut_m_summa$effort))



```


# export table S1

```{r t1, echo=FALSE, verbose = FALSE}

knitr::kable(s.sizes.b,caption = "Table S1", digits =2) %>% kable_styling()# %>%save_kable(here("tableS1_sampleSizes_bic.html"))
knitr::kable(s.sizes.m,caption = "Table S1", digits =2) %>% kable_styling() #%>%save_kable(here("tableS1_sampleSizes_metis.html"))

```

# main results - abundance 
```{r ts3}
N_bic.3_summa <- N_bic.3_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(N_bic.3_summa,caption = "Harbour seal pup abundance in the St. Lawrence Estuary, 1998-2019, at Bic Island, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.", "simple")
```


```{r}
N_metis.3_summa <- N_metis.3_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(N_metis.3_summa,caption = "Harbour seal pup abundance in the St. Lawrence Estuary, 1998-2019, at Métis, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.", "simple")
```


# main results - surv 
```{r}
weanSurvOut_b_summa <- weanSurvOut_b_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(weanSurvOut_b_summa,caption = "Harbour seal pup pre-weaning survival in the St. Lawrence Estuary, 1998-2019, at Bic Island, Québec, Canada. Estimates are from a model with constant (mean) intra-annual survival and fixed time effects on detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.") # %>%save_kable(here("tableS7_surv_bic.html"))
```

```{r}
weanSurvOut_m_summa <- weanSurvOut_m_summa %>% mutate(across(where(is.numeric), round, 2))
knitr::kable(weanSurvOut_m_summa,caption = "Harbour seal pup abundance in the St. Lawrence Estuary, 1998-2019, at Métis, Québec, Canada. Estimates are from a model with random time effects on intra-annual survival and fixed time effects on detection probability. Mean posterior estimates, 25th and 75th quantiles, and 95% credible intervals around posterior means were extracted. Data were not available for years 2004-2007 and 2017-2018.") # %>%save_kable(here("tableS7_surv_metis.html"))

```


These outputs can be compared - one includes medians, the other only means - otherwise outputs are equivalent.
```{r abundance and survival}
results_jollySeber <- results_jollySeber %>% mutate(across(where(is.numeric), round, 2))
# results_weanSurv <- results_weanSurv %>% mutate(across(where(is.numeric), round, 2))




# compare to gab's extract
weanSurvOut_m <-  weanSurvOut_m%>% mutate(across(where(is.numeric), round, 2))
weanSurvOut_m_summa <-  weanSurvOut_m_summa%>% mutate(across(where(is.numeric), round, 2))

knitr::kable(weanSurvOut_m_summa,caption = "Table S1", digits =2) %>% kable_styling()# %>%save_kable(here("tableS1_sampleSizes_bic.html"))


weanSurvOut_b <- weanSurvOut_b%>% mutate(across(where(is.numeric), round, 2))
weanSurvOut_b_summa <-  weanSurvOut_b_summa%>% mutate(across(where(is.numeric), round, 2))

knitr::kable(weanSurvOut_b_summa,caption = "Table S1", digits =2) %>% kable_styling()# %>%save_kable(here("tableS1_sampleSizes_bic.html"))


# annual trend?
summary(lm(weanSurv_mean~yr, data=weanSurvOut_m_summa)) # yes at METIS
# yr          -0.015594   0.005975  -2.610   0.0206 *
summary(lm(weanSurv_mean~as.numeric(yr), data=weanSurvOut_b_summa)) # no at BIC
# as.numeric(yr) -0.006997   0.009975  -0.701    0.495
```



Get the birthdates but do not focus on that given high ci. 
```{r}

# birth date 
sapply(outlist_b,function(x) x$summary$all.chains['mu.bd',]) 
meanBd_bic <- sapply(outlist_b,function(x) x$summary$all.chains['mu.bd',]) %>% 
    t() %>% as.data.frame() %>% rename(CIL=`95%CI_low`,CIH=`95%CI_upp`) %>% mutate(yr=years)

sapply(outlist_m,function(x) x$summary$all.chains['mu.bd',]) 


meanBd_metis <- sapply(outlist_m,function(x) x$summary$all.chains['mu.bd',]) %>% 
  t() %>% as.data.frame() %>% rename(CIL=`95%CI_low`,CIH=`95%CI_upp`) %>% mutate(yr=years)

# here the birth dates as estimated from the models
meanBd_bic <-  meanBd_bic %>% mutate(across(where(is.numeric), round, 1))
meanBd_metis <-  meanBd_metis %>% mutate(across(where(is.numeric), round, 1))
```


# correlations between surv and n


# attention years with low convergence can not be used. 



```{r}
correlation_bic<- lbda_bic.3 %>% filter(yr!=2002&yr!=2008) %>% group_by(it) %>% 
    summarise(cor_sN=cor(weanSurv,Nsuper),
              cor_sgrowth=cor(weanSurv,growth))
correlation_Metis<- lbda_metis.3 %>% filter(yr!=2000&yr!=2001&yr!=2002&yr!=2013) %>% group_by(it) %>% 
  summarise(cor_sN=cor(weanSurv,Nsuper),
            cor_sgrowth=cor(weanSurv,growth))



# median correlation between pre-weaning survival and pup abundance 
quantile(correlation_bic$cor_sN,probs = c(0.025,.5,0.975))# -0.44395513  0.05559576  0.52749523 
quantile(correlation_Metis$cor_sN,probs = c(0.025,.5,0.975)) # -0.6030714 -0.2743276  0.3713584 

#  growth
quantile(correlation_bic$cor_sgrowth,probs = c(0.025,.5,0.975)) #-0.592102299 -0.001278801  0.627859854 
round(quantile(correlation_Metis$cor_sgrowth,probs = c(0.025,.5,0.975)),2) # -0.44  0.02  0.53


hist(correlation_bic$cor_sN);abline(v=0,col='red')
mean(correlation_bic$cor_sN>0)  # pseudo p-value  0.5113333 
  
hist(correlation_bic$cor_sgrowth);abline(v=0,col='red')
mean(correlation_bic$cor_sgrowth>0)  # pseudo p-value 0.721

hist(correlation_Metis$cor_sN);abline(v=0,col='red')
mean(correlation_Metis$cor_sN>0)  # pseudo p-value 0.1086667

hist(correlation_Metis$cor_sgrowth);abline(v=0,col='red')
mean(correlation_Metis$cor_sgrowth>0)  # pseudo p-value = 0.696


# mean growth
mean(correlation_Metis$growth)

```

# get mean lambda 
```{r}
 t= lbda_bic.3 %>% group_by(yr) %>%
    summarise(lbda_mean=mean(growth),
              lbda_cil=as.numeric(quantile(growth,0.025)),
              lbda_c25=as.numeric(quantile(growth,0.25)),
              lbda_c75=as.numeric(quantile(growth,0.75)),
              lbda_cih=as.numeric(quantile(growth,0.975))
    )
mean(t$lbda_mean) #1.160159
  
tm=lbda_metis.3 %>% group_by(yr) %>%
    summarise(lbda_mean=mean(growth),
              lbda_cil=as.numeric(quantile(growth,0.025)),
              lbda_c25=as.numeric(quantile(growth,0.25)),
              lbda_c75=as.numeric(quantile(growth,0.75)),
              lbda_cih=as.numeric(quantile(growth,0.975))
    )
mean(tm$lbda_mean)
  
survb=lbda_bic.3 %>% group_by(it) %>% filter(yr!=2008&yr!=2002) %>% 
    summarise(meanSurv=mean(weanSurv))
quantile(survb$meanSurv,probs = c(0.025,.5,0.975)) # 
#      2.5%       50%     97.5% 
# 0.4449693 0.5684772 0.6830830 
mean(survb$meanSurv)     # 0.5681413


survm=lbda_metis.3 %>% group_by(it) %>% filter(yr!=2000&yr!=2001&yr!=2002&yr!=2013) %>% 
    summarise(meanSurv=mean(weanSurv))
quantile(survm$meanSurv,probs = c(0.025,.5,0.975)) # 
#      2.5%       50%     97.5% 
# 0.1750356 0.2973254 0.4347385 
     
mean(survm$meanSurv)  # 0.2993452


```

```{r}

```


Finally, we investigated if pre-weaning survival was correlated either with pup abundance or the 'observed population growth rate', calculated as Nt+1/Nt, where N was pup abundance. Pre-weaning survival was not correlated with neither pup abundance at time t nor pup growth rate at Bic (median correlation with N [95% CRI]: `r  round(quantile(correlation_bic$cor_sN,probs = c(.5)), 2) ` [`r  round(quantile(correlation_bic$cor_sN,probs = c(.025)), 2) `, `r  round(quantile(correlation_bic$cor_sN,probs = c(.975)), 2) `], median correlation with growth rate [95% CRI]: `r  round(quantile(correlation_bic$cor_sgrowth,probs = c(.5)), 2) ` [`r  round(quantile(correlation_bic$cor_sgrowth,probs = c(.025)), 2) `, `r  round(quantile(correlation_bic$cor_sgrowth,probs = c(.975)), 2) `] or at Metis (median correlation with N [95% CRI]: `r  round(quantile(correlation_Metis$cor_sN,probs = c(.5)), 2) ` [`r  round(quantile(correlation_Metis$cor_sN,probs = c(.025)), 2) `, `r  round(quantile(correlation_Metis$cor_sN,probs = c(.975)), 2) `], median correlation with growth rate [95% CRI]: `r  round(quantile(correlation_Metis$cor_sgrowth,probs = c(.5)), 2) ` [`r  round(quantile(correlation_Metis$cor_sgrowth,probs = c(.025)), 2) `, `r  round(quantile(correlation_Metis$cor_sgrowth,probs = c(.975)), 2) `]. 



# Tables and figures{-}

Here figures are for model 3 in both cases but will have to check convergence issues first. 
```{r}
mytheme <- theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"), legend.position = "none",
                 axis.text=element_text(size=12, color="white"),
                 axis.title=element_text(size=14, color="white"),
                 axis.ticks.x =element_line(color="white"),
                 axis.ticks.y = element_line(color="white"),
                 rect = element_rect(fill = "white"),
                 axis.line = element_line(color="white"),
                 panel.grid.major = element_line(color="white"),
                 panel.grid.minor = element_line(color="transparent"),
                 panel.background = element_rect(fill = "transparent",colour = NA),
                 plot.background = element_rect(fill = "transparent",colour = NA))
background_grid(
    major = c("xy", "x", "y", "only_minor", "none"),
    minor = c("none", "xy", "x", "y"),
    size.major = 0.5,
    size.minor = 0.2,
    color.major = "grey85",
    color.minor = "grey85"
)

```
# figure 1 - map of study area

# figure 2 - abundance at low and high capture rates 

```{r f3, echo=FALSE, fig.cap="Capture probability over time in harbour seal pups, Bic Island, Quebec, Canada, in a year with a) low sample size and recapture rate (2019, N = 56, recapture rate 0.05 and b) high sample size and recapture rate (year 1999, N=37, recapture rate = 0.81). full description of sample size per year appears in Table 1. ", fig.height=5, fig.width=7.165354}
# cowplot::plot_grid(p_panel[[16]],p_panel[[2]], labels=c('a)','b)'))
```

# figure 3 - abundance over years 

```{r f1, echo=FALSE, fig.cap="Harbour seal pup abundance in the St. Lawrence estuary, 1998-2019, at Bic Island and Métis, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Bars represent 95% credible intervals around posterior mean. Data were not available for years 2004-2007 and 2017-2018.", fig.width=7.165354, fig.height=5}
# results_jollySeber %>%
#   ggplot(aes(x=yr, y=N, ymin=ymin, ymax=ymax, color=site)) +
#   geom_pointrange(position=position_dodge(w=0.5), size=0.75) +
#   labs(x=expression('Year')) +
#   labs(y="Pup abundance") +
#  # scale_y_continuous(limits = c(0,1),breaks = seq(from = 0, to = 1, by = 0.1)) +
#   scale_x_continuous(breaks = seq(from = 1998, to = 2019, by = 2)) +
#   #mytheme + background_grid()
#   theme(panel.grid.minor = element_line(color="transparent"))+
#   scale_colour_manual(values = c("bic" = "#FFDB6D", "metis" = "#00AFBB")) +
#   theme_cowplot(12) +
#   theme(panel.grid.major = element_line(color="lightgrey", linetype = 'dotted'))


N_bic.3_summa$site <- 'bic'
N_metis.3_summa$site <- 'metis'
df_N <- rbind(N_bic.3_summa, N_metis.3_summa)
knitr::kable(df_N,caption = "Table S5", digits =1) %>% kable_styling() # %>%  save_kable(here("tableS5_abundance.html"))


ggplot(df_N,aes(x=as.factor(yr), color = site))+
    geom_linerange(aes(ymin=Nsuper_cil,ymax=Nsuper_cih), position=position_dodge(w=0.75), size=0.75)+
    geom_linerange(aes(ymin=Nsuper_c25,ymax=Nsuper_c75),size=2, position=position_dodge(w=0.75), size=0.75)+
    geom_point(aes(y=Nsuper_median),shape=18,size=4, position=position_dodge(w=0.75), size=0.75)+
    labs(x='Year',y='Pup abundance')+
  scale_y_continuous(limits = c(0,400),breaks = seq(from = 0, to = 400, by = 50)) +
  theme(panel.grid.minor = element_line(color="transparent"))+
  scale_colour_manual(labels = c("Bic Island", "Métis"), values = c("bic" = "#FFDB6D", "metis" = "#00AFBB")) + theme_cowplot(10) +
  theme(panel.grid.major = element_line(color="lightgrey", linetype = 'dotted'))

#ggsave("FIGS3_abundanceMedian.png", width=8, height=4.5,unit='in', dpi = 600)

```
# figure 4 - survival over years 

```{r f2, echo=FALSE, fig.cap="Pre-weaning survival of harbour seal in the St. Lawrence estuary, 1998-2019, at Bic Island and Métis, Québec, Canada. Estimates are from a model with a fixed effect of sampling occasions on intra-annual survival and detection probability. Bars represent 95% credible intervals around posterior mean. Data were not available for years 2004-2007 and 2017-2018. Two years did not converge: 2003 for Bic Island, and 2013 for Métis; the estimates are thus excluded from the figure.", fig.height=5, fig.width=7.165354}

# THIS IS VERY CONSERVATIVE - I EXCLUDE 3 YEARS WITH CONVERGENCE PROBLEMS 

# get means (june version)
weanSurvOut_b_summa_filtered <- weanSurvOut_b_summa %>% filter(yr!=2008 & yr!=2002)
mean(weanSurvOut_b_summa_filtered$weanSurv_mean) #  0.4613333
quantile(weanSurvOut_b_summa_filtered$weanSurv_mean,probs = c(0.025,.5,0.975)) 
#  2.5%    50%  97.5% 
# 0.0235 0.4700 0.7730 

weanSurvOut_m_summa_filtered <- weanSurvOut_m_summa %>% filter(yr!=2000&yr!=2002&yr!=2013)# this will need to change depending on final models 
mean(weanSurvOut_m_summa_filtered$weanSurv_mean) #   0.2628571 much lower than before 
quantile(weanSurvOut_m_summa_filtered$weanSurv_mean,probs = c(0.025,.5,0.975)) 
# 2.5%   50% 97.5% 
# 0.02975 0.22500 0.63775 

# try wit this df 
weanSurvOut_b_summa_filtered$site <- 'bic'
weanSurvOut_m_summa_filtered$site <- 'metis'
df_weanSurv <- rbind(weanSurvOut_b_summa_filtered, weanSurvOut_m_summa_filtered)
# df_weanSurv <- df_weanSurv %>% filter(yr!=2003|site!='bic') %>% filter(yr!=2013|site!='metis')

knitr::kable(df_weanSurv,caption = "Table S7", digits =1) %>% kable_styling() # %>% save_kable(here("tableS7_survival.html"))


# figure 2022 09 09 now with median 
ggplot(df_weanSurv,aes(x=as.factor(yr), color = site))+
    geom_linerange(aes(ymin=weanSurv_cil,ymax=weanSurv_cih), position=position_dodge(w=0.75), size=0.75)+
    geom_linerange(aes(ymin=weanSurv_c25,ymax=weanSurv_c75),size=2, position=position_dodge(w=0.75), size=0.75)+
    geom_point(aes(y=weanSurv_med),shape=18,size=4, position=position_dodge(w=0.75), size=0.75)+
    labs(x='Year',y='Pre-weaning survival')+
    scale_y_continuous(limits = c(0,1),breaks = seq(from = 0, to = 1, by = 0.1)) +
  theme(panel.grid.minor = element_line(color="transparent"))+
  scale_colour_manual(labels = c("Bic Island", "Métis"),values = c("bic" = "#FFDB6D", "metis" = "#00AFBB")) + theme_cowplot(10) +
  theme(panel.grid.major = element_line(color="lightgrey", linetype = 'dotted')) 

#ggsave("FIG4_survivalMedian_2022-12-09.png", width=8, height=4.5,unit='in', dpi = 600)

```



# additional analyses with survival
```{r}

```





# supplementary tables and figures 

## parturition date from CJS at bic 

```{r echo=FALSE, verbose=FALSE}
# reorganise a bit 
colnames(meanBd_bic)
# # meanBd_bic <- meanBd_bic[, c("yr" , "Mean","Median","St.Dev.", "CIL","CIH")]
# names(meanBd_bic) <- c("Year" , "MeanDate","MedianDate","St.Dev.", "Lower CRI","Upper CRI")

knitr::kable(meanBd_bic,caption = "Estimated parturition date of harbour seal pups, Bic Islanc, Quebec, Canada, 1998-2019. Parturition date was modeled as a function of body mass within a Cormack-Jolly-Seber model of pre-weaning survival. Each year was modeled independently and estimates are from a model with a fixed time effect on capture probability and a constant daily survival probability over time. No data were available for 2004-2007 and 2017-2018. CRI: 95% credibilibity interval.") # %>% save_kable('TABLES8_birthdates_bic.html')
```
## parturition date from CJS at metis 

```{r echo=FALSE, verbose=FALSE}
# reorganise a bit 
colnames(meanBd_metis)
# meanBd_metis <- meanBd_metis[, c("yr" , "Mean","Median","St.Dev.", "CIL","CIH")]
# names(meanBd_metis) <- c("Year" , "Mean date","Median date","St.Dev.", "Lower CRI","Upper CRI")

knitr::kable(meanBd_metis,caption = "Estimated parturition date of harbour seal pups, Metis, Quebec, Canada, 1998-2019. Parturition date was modeled as a function of body mass within a Cormack-Jolly-Seber model of pre-weaning survival. Each year was modeled independently and estimates are from a model with a fixed time effect on capture probability and a random time effect on daily survival probability. No data were available for 2004-2007 and 2017-2018. CRI: 95% credibilibity interval.") # %>% save_kable('TABLES8_birthdates_metis.html')
```


